\section{Method} \label{Methoden} 
\subsection{Archives and Search Queries}
\textit{PubMed}, \textit{Science direct}, \textit{PsycInfo} were used as search archives. An overall scope of the topic was made with search terms "machine learning" <OR> "deep learning" <AND> "mri" <OR> "neuroimaging" <AND> "alzheimer" <AND> "review" <OR> "systematic". After this process the focus of this review was shifted toward the two model \gls{svm} and \gls{cnn}. Therefore, two different groups of search terms were applied to the search engine. The first group of search terms was used to find relevant studies that used \gls{svm} model. Search terms consisted of "svm" <AND> "mri" <AND> "alzheimer". The second group of search terms was used to find papers focusing on \gls{cnn} and consisted of "cnn" <AND> "mri" <AND> "alzheimer". The search was applied to title and abstract section of the paper. Furthermore, the search was limited to studies in English and from year 2020 and 2021. 

\subsection{Study Inclusion Criteria}
Applying the first group of search terms yielded 14 results on \textit{PubMed}. Studies using fMRI or focusing heavily on MCI, or not focus on \gls{AD} classification were excluded. Results were compared to the same search queries in \textit{Science direct} and \textit{PsycInfo}, finally, arriving at five studies using SVM. 
Using the second group of search terms resulted in 30 papers on \textit{PubMed}. After cross referencing with other search archives, 17 papers passed inclusion criteria. Since not all 17 studies could be reviewed, further inclusion considerations were made. 
These considerations were: the study should only use MRI as modality and have a sample size higher than 400 . The sample size is based on reflections made by \autocite{ebrahimighahnaviehDeepLearningDetect2020} that used sample size to rate the quality of their included studies. Finally, five studies were chosen for further review. Therefore, five studies using SVM and five studies using CNN amounting to a total of 10 studies were included in this review. 